---
title: "EDA_FA4_KHAFAJI"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(dcldata)
library(MASS)

```

# Exploring the Mortality by Latitude dataset

First, let's load the dataset.

```{r read mortality by lat}
mbl <- read.csv("mortality_by_latitude.csv", sep=",")
mbl

```

### 1 Make a plot of the mortality index against mean average temperature.  

```{r mortality index vs mean average temp}
mbl %>% ggplot(aes(y = mortality_index, x=temperature)) +
  geom_point() +
  geom_smooth(method=loess
              , se=FALSE)

```

From here, we can conclude that the data is hollow up. Let's find a transformation that can help with that.

```{r transformation}

mbl <- mbl %>% mutate(temperature_5thPower = temperature^5)

mbl %>% ggplot(aes(y = mortality_index, x=temperature_5thPower)) +
  geom_point() +
  geom_smooth(method=loess
              , se=FALSE)+
  labs(title="Temperature^5 vs Mortality Index", x="Temperature^5", y="Mortality Index")
```

We can see that 

let's then make a plot of the residuals to check for any remaining patterns.

```{r residual plot}
model <- lm(mortality_index ~ temperature_5thPower, data=mbl)

model %>% ggplot(aes(x=.fitted, y=.resid)) +
  geom_point()+
  geom_hline(yintercept=0)+
  labs(title="Residual Plot of Model trained on Temperature^5 vs Mortality Index", x="Fitted Values", y="Residuals")

```

We can see that The model is not that accurate with low temperatures, but increases in accuracy as temperature increases.

# Exploring the Diamonds Dataset

## 2. Using the same subset of the diamonds dataset, make a plot of log price as a function of carat with a loess smoother. Try several values for the span and degree arguments and comment briefly about your choice.

```{r dias 2}
diamonds %>% ggplot(aes(x=carat, y=log(price)))+
  geom_point(alpha=0.2)+
  geom_smooth(method = "loess", se = FALSE, span = 0.3, method.args = list(degree = 1), color = "blue") +
  geom_smooth(method = "loess", se = FALSE, span = 0.5, method.args = list(degree = 1), color = "red") +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, method.args = list(degree = 2), color = "green") +
  geom_smooth(method = "loess", se = FALSE, span = 0.5, method.args = list(degree = 2), color = "yellow")
```


The loess fit with a span of 0.5 and degree = 2 fit the data better than the other spans and degrees


## 3. Compare the fit of the loess smoother to the fit of the polynomial + step function regression using a plot of the residuals in the two models. Which one is more faithful to the data? 

```{r residuals of models}

diamonds <- diamonds %>% mutate(log_price = log(price))

loess_fit <- loess(log_price ~ carat, data = diamonds, span = 0.5, degree = 2)
diamonds <- diamonds %>%
  mutate(loess_resid = residuals(loess_fit))


diamonds <- diamonds %>%
  mutate(carat_bin = cut(carat, breaks = seq(0, 10, by = 0.25)))
poly_step_fit <- lm(log_price ~ poly(carat, degree = 2) + carat_bin, data = diamonds)
diamonds <- diamonds %>%
  mutate(poly_step_resid = residuals(poly_step_fit))

ggplot(diamonds) +
  geom_point(aes(x = carat, y = loess_resid, color = "LOESS Residuals"), alpha = 0.3) +
  geom_point(aes(x = carat, y = poly_step_resid, color = "Polynomial + Step Residuals"), alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals Comparison: LOESS vs. Polynomial + Step Function",
       x = "Carat",
       y = "Residuals",
       color = "Model") +
  theme_minimal()
```

As we can see, although they have almost the same distribution, the Polynomial + Step regression has a smaller absolute value of the residuals. Therefore, Polynomial + Step regression is the better model for log price as a function of carat.













